A load balancing scheme for massively multiplayer online games \thanks{This work was supported by the National Research Council (CNPq) and by the Coordination of Improvement of Higher Education (CAPES), both brazilian research funding agencies.}

### 2009-04-10 trunk/textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

To perform the simulation of the load balancing, a virtual environment was simulated with various avatars in it moving according to the random waypoint model \cite{bettstetter2004spr}. However, the choice of which waypoint go was not completely random. Three hotspots were defined, and there was a probability for the havatar choose one of these hotspots as its next waypoint. The purpose of this was to force an uneven distribution of avatars in the virtual environment, putting to test the algorithms of the proposed phase 2. Those who take account of the existence of hotspots form the regions on this basis, reducing the distribution overhead.

The environment consisted of a two-dimensional space, divided into 225 cells, forming a matrix of order 15. There were 750 avatars. All cells always belonged to some region, although they could be transferred between regions. There were eight servers ($S_1, S_2, ..., S_8$), each one of them associated to a region, each of which could have 0 to all 225 cells. The capacity of all servers was different, with $P(S_i) = i \times 20000$. Thus, it was possible to test whether the distribution obeyed the criterion of proportional load balancing. Each session of simulated game was 20 minutes long. The total weight of the game was purposely set in a way that it was greater than the total capacity of the system, forcing the triggering of the load balancing.

>>>>

To perform the simulation of the load balancing, a virtual environment was simulated with various avatars in it moving according to the random waypoint model \cite{bettstetter2004spr}. However, the choice of which waypoint to go was not completely random. Three hotspots were defined, and there was a probability for the avatar to choose one of these hotspots as its next waypoint. The purpose of this was to force an uneven distribution of avatars in the virtual environment, putting to test the load balancing algorithms. Those who take account of the existence of hotspots form the regions on this basis, reducing the distribution overhead.

As we said in section \ref{sec:related}, we also compare our algorithms to the one proposed by Ahmed and Shirmohammadi \cite{ahmed2008mol}. However, it is very important to observe that some changes have been made before simulating. First, the authors divide the virtual environment in hexagonal cells, and we simulate their algorithm on a grid-like division, with square cells. Moreover, as we consider an heterogeneous system, each server $S_i$ presents its own message rate threshold, $T_i^m$.
%
Nonetheless, we keep what we consider as the core of their approach, which is the selection of the smallest cell cluster managed by the overloaded server, followed by the selection of the cell of that cluster which has the lowest interaction with other cells of the same cluster and, finally, the transference of the selected cell to the least loaded server.

The environment consisted of a two-dimensional space, divided into 225 cells, forming a matrix of order 15. There were 750 avatars. Each cell always belonged to some region, although it could be transferred to another region. There were eight servers ($S_1, S_2, ..., S_8$), each one of them associated to a region, each of which could have 0 to all 225 cells. The capacity of all servers was different, with $P(S_i) = i \times 20000$. Thus, it was possible to test whether the distribution obeyed the criterion of proportional load balancing. Each session of simulated game was 20 minutes long. The total weight of the game was purposely set in a way that it was greater than the total capacity of the system, forcing the triggering of the load balancing.
